---
title: "Assignment 1 - Language Development in ASD - part 3"
author: "Riccardo Fusaroli"
date: "August 10, 2018"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
getwd()
library(pacman)
p_load(lmerTest, pastecs, ggplot2, tidyverse, gdata, MuMIn, effects, stringr, plyr, vtreat, cvTools, caret)
```

## Welcome to the third exciting part of the Language Development in ASD exercise

In this exercise we will delve more in depth with different practices of model comparison and model selection, by first evaluating your models from last time, then learning how to cross-validate models and finally how to systematically compare models.

N.B. There are several datasets for this exercise, so pay attention to which one you are using!

1. The (training) dataset from last time (the awesome one you produced :-) ).

2. The (test) datasets on which you can test the models from last time:

* Demographic and clinical data: https://www.dropbox.com/s/ra99bdvm6fzay3g/demo_test.csv?dl=1
* Utterance Length data: https://www.dropbox.com/s/uxtqqzl18nwxowq/LU_test.csv?dl=1
* Word data: https://www.dropbox.com/s/1ces4hv8kh0stov/token_test.csv?dl=1

```{r, include=FALSE}
#Loading the new data 
demo_test <- read.csv("demo_test.csv")
token_test <- read.csv("token_test.csv")
lu_test <- read.csv("lu_test.csv")
```

```{r, include=FALSE}
#Renaming the data sets to have the same headline for the subject = ID
demo_test <- rename.vars(demo_test, "Child.ID", "ID")
lu_test <- rename.vars(lu_test, "SUBJ", "ID")
token_test <- rename.vars(token_test, "SUBJ", "ID")

#Renaming the visit to have the same 
lu_test <- rename.vars(lu_test, "VISIT", "visit")
token_test <- rename.vars(token_test, "VISIT", "visit")
demo_test <- rename.vars(demo_test, "Visit", "visit")

#Homogeneize the way visit is reported, use of stringr to extract only numbers
lu_test$visit <- str_extract(lu_test$visit, "[1-6]")
token_test$visit <- str_extract(lu_test$visit, "[1-6]")

#Removing all punctuations in the ID column 
demo_test$ID <- str_replace_all(demo_test$ID, "[:punct:]", "")
lu_test$ID <- str_replace_all(lu_test$ID, "[:punct:]", "")
token_test$ID <- str_replace_all(token_test$ID, "[:punct:]", "")

#Selecting the variables needed
demo_test_sub <- select(demo_test, c(ID, visit, Ethnicity, Diagnosis, Gender, Age, ADOS,  MullenRaw, ExpressiveLangRaw))
lu_test_sub <- select(lu_test, c(ID, visit, MOT_MLU, MOT_LUstd, CHI_MLU, CHI_LUstd))
token_test_sub <- select(token_test, c(ID, visit, types_MOT, types_CHI, tokens_MOT, tokens_CHI))

library(plyr)
#Using join to merge to three datasets by ID and Visit
test <- join(demo_test_sub, lu_test_sub)
test <- join(test, token_test_sub)

#Ados for only visit 1
#Using subset to make a dataset only consisting of the data from visit 1
subset_test <- subset(test, visit == 1, select = c(ID, ADOS, MullenRaw, ExpressiveLangRaw))

#Changing the names in the new data set 
subset_test <- rename.vars(subset_test, c("ADOS", "MullenRaw", "ExpressiveLangRaw"), c("ADOS1", "NonverbalIQ", "VerbalIQ"))

#Merging the two datasets using join 
test <- join(test, subset_test)

#Using select to make a clean dataset with variables in the right order
clean_test <- select(test, c(ID, Diagnosis, visit, Gender, Age, Ethnicity,  ADOS1, NonverbalIQ, VerbalIQ, CHI_MLU, MOT_MLU, CHI_LUstd, MOT_LUstd, types_CHI, types_MOT,  tokens_CHI, tokens_MOT))

#Anonymous kids
clean_test$ID <- as.factor(clean_test$ID)
clean_test$ID <- as.numeric(clean_test$ID)

#Turning the 1s and 2s into M and F in the gender variable
clean_test$Gender <- ifelse(clean_test$Gender == "1", "M", "F")

#Turning the As and Bs into ADS and TD in the diagnosis variable using ifelse (short way)
clean_test$Diagnosis <- ifelse(clean_test$Diagnosis == "A", "ASD", "TD")

#Making a csv. file named LanguageASD.csv to WD
write.csv(clean_test, file = "test_data_clean_v1.csv")

```

#######Start loading from here#######
Loading CSV's

```{r, include=FALSE}
#Loading data (if cleared workspace)
data_test <- read.csv("test_data_clean_v1.csv")

#Loading old data
data = read.csv("LanguageASD_final.csv")
```

Changing names in the traning data
```{r,include=FALSE}
data <- rename.vars(data, "Diagnosis", "diagnosis")
data <- rename.vars(data, "Gender", "gender")
data <- rename.vars(data, "Age", "age")
data <- rename.vars(data, "Ethnicity", "ethnicity")
data <- rename.vars(data, "Visit", "visit")
data <- rename.vars(data, "ADOS1", "ados1")
data <- rename.vars(data, "MullenRaw1", "nonverbalIQ")
data <- rename.vars(data, "ExpressiveLangRaw1", "verbalIQ")

data <- data[-c (1, 68, 69, 130, 131, 132, 276, 277, 284, 285), ]
```

Changing names in the test data
```{r, include=FALSE}
data_test <- rename.vars(data_test, "Diagnosis", "diagnosis")
data_test <- rename.vars(data_test, "Gender", "gender")
data_test <- rename.vars(data_test, "Age", "age")
data_test <- rename.vars(data_test, "Ethnicity", "ethnicity")
data_test <- rename.vars(data_test, "ADOS1", "ados1")
data_test <- rename.vars(data_test, "NonverbalIQ", "nonverbalIQ")
data_test <- rename.vars(data_test, "VerbalIQ", "verbalIQ")
```

### Exercise 1) Testing model performance

How did your models from last time perform? In this exercise you have to compare the results on the training data () and on the test data. Report both of them. Compare them. Discuss why they are different.

- recreate the models you chose last time (just write the model code again and apply it to your training data (from the first assignment))
- calculate performance of the model on the training data: root mean square error is a good measure. (Tip: google the function rmse())
- create the test dataset (apply the code from assignment 1 part 1 to clean up the 3 test datasets)
- test the performance of the models on the test data (Tips: google the functions "predict()")
- optional: predictions are never certain, can you identify the uncertainty of the predictions? (e.g. google predictinterval())

formatting tip: If you write code in this document and plan to hand it in, remember to put include=FALSE in the code chunks before handing in.

```{r, include=FALSE}
#data$CHI_MLU <- data$CHI_MLU[!is.na(data$CHI_MLU)]
p_load(Metrics)

#Removing NAs in data sets
data <- na.omit(data, cols = "CHI_MLU")
data_test <- na.omit(data_test, cols = "CHI_MLU")

#Using our best model from the last assignment
model_train <- lmer(CHI_MLU ~ visit + I(visit^2) + ados1 + MOT_MLU + verbalIQ + (1+visit + I(visit^2) | ID), data = data, REML=F)

#Root mean square error on how the model perform on the training data
rmse(data$CHI_MLU, predict(model_train))
##[1] 0.2775267

#Very bad prediction on the test set 
rmse(data_test$CHI_MLU, predict(model_train, data_test))
##[1] 0.7683667
```

#### ANSWER: 
Our selected model from last assignment (pseudocode: mean length of utterance ~ visit + visit^2 + ados + mother_mlu + (visit + visit^2 | ID)) produces a root mean square error of 0.278. The root mean square error increases quite a lot when applying the model on the test set, from 0.278 to 0.768. We thus observe that the accuracy of the predictions decreases a lot when the model is applied to the new (test) data, indicating that the model's predictive power is quite low. As such, this model seem to contain a fair amount explanatory power, but the predictive power is low.

### Exercise 2) Model Selection via Cross-validation (N.B: ChildMLU!)

One way to reduce bad surprises when testing a model on new data is to train the model via cross-validation.
In this exercise you have to use cross-validation to calculate the predictive error of your models and use this predictive error to select the best possible model.

- Use cross-validation to compare your model from last week with the basic model (Child MLU as a function of Time and Diagnosis, and don't forget the random effects!)

- (Tips): google the function "createFolds";  loop through each fold, train both models on the other folds and test them on the fold)

Which model is better at predicting new data: the one you selected last week or the one chosen via cross-validation this week?

- Test both of them on the test data.
- Report the results and comment on them.
- Now try to find the best possible predictive model of ChildMLU, that is, the one that produces the best cross-validated results.

```{r, include=FALSE}
#Creating the basic model
hercules <- lmer(CHI_MLU ~ visit + diagnosis + (1 + visit | ID), data = data, REML = F)

#Creating our model from last week
zeus <- lmer(CHI_MLU ~ visit + I(visit^2) + ados1 + MOT_MLU + verbalIQ + (1+visit + I(visit^2)| ID), data = data, REML=F)
```

```{r}
#Cross-validation of the basic model
set.seed(242)
folds <- createFolds(unique(data$ID),5)

n=1
rmsetrain = NULL
rmsetest = NULL

for (f in folds){
  train_data <- subset(data, !(ID %in% f))
  test_data <- subset(data, ID %in% f)
  model <- lmer(CHI_MLU ~ visit + diagnosis + (1 + visit| ID), data = train_data, REML = FALSE)
  rmsetrain[n] = rmse(train_data$CHI_MLU, predict(model))
  rmsetest[n] = rmse(test_data$CHI_MLU, predict(model, test_data, allow.new.levels = TRUE))
  n = n +1
}

# RMSE of model prediction when fitted to the full dataset
rmse(data$CHI_MLU, predict(hercules))
#[1] 0.3380725

# RMSE of model prediction using cross-validation
rmse_basic <- data.frame(rmsetest, rmsetrain)
mean(rmse_basic$rmsetest)
#[1] 0.8116268

#Cross-validation of our model
n=1
rmsetrain = NULL
rmsetest = NULL

for (f in folds){
  train_data <- subset(data, !(ID %in% f))
  test_data <- subset(data, ID %in% f)
  model <- lmer(CHI_MLU ~ visit + I(visit^2) + ados1 + MOT_MLU + verbalIQ + (1 + visit + I(visit^2) | ID), data = train_data, REML=FALSE)
  rmsetrain[n] = rmse(train_data$CHI_MLU, predict(model))
  rmsetest[n] = rmse(test_data$CHI_MLU, predict(model, test_data, allow.new.levels = TRUE))
  n = n +1
}

# RMSE of model prediction when fitted to the full dataset
rmse(data$CHI_MLU, predict(zeus))
#[1] 0.2775267

# RMSE of model prediction using cross-validation
rmse_fav <- data.frame(rmsetest, rmsetrain)
mean(rmse_fav$rmsetest)
#[1] 0.6341372
```

```{r, include=FALSE}
# Model predictions when using the separate test-data
rmse(predict(hercules, data_test), data_test$CHI_MLU)
#[1] 1.068573

rmse(predict(zeus, data_test), data_test$CHI_MLU)
#[1] 0.7683667
```

#### ANSWER: 
The basic linear model (pseudocode: mean length of utterance ~ visit + diagnosis + (visit | ID)) produces a root mean squared error of only 0.34 when fitted to the full dataset (test data excluded). However, when the model is trained and tested using 5-fold cross validation, the root mean squared error increases to 0.79.

As previously mentioned, our selected model (pseudocode: mean length of utterance ~ visit + visit^2 + verbalIQ + ados + mother's MLU (visit + visit^2 | ID)) produces a root mean squared error of only 0.28 when fitted to the full dataset (test data excluded). As with the linear model, the error increases when the model is trained and tested using 5-fold cross validation, in this case to roughly 0.62.

First of all, this means that when cross-validation our selected model remains superior in making predictions compared to the linear model.

When the two model predictions are applied to the actual test data they both turn out to be quite far off. The linear model produces a root mean squared error of 1.07 while the quadratic model produces a RMSE of 0.77. Thus, the quadratic model is still performing better on the test data than the linear, however somewhat worse than predicted with the training data (both with and without cross validation of the model, but especially the former).

```{r}
#Testing Nanna's random model
n=1
rmsetrain = NULL
rmsetest = NULL

for (f in folds){
  train_data <- subset(data, !(ID %in% f))
  test_data <- subset(data, ID %in% f)
  model_curls <- lmer(CHI_MLU ~ visit + I(visit^2) + ados1 + verbalIQ + MOT_MLU + types_CHI + (1+visit + I(visit^2)| ID), data = train_data, REML = FALSE)
  rmsetrain[n] = rmse(train_data$CHI_MLU, predict(model_curls))
  rmsetest[n] = rmse(test_data$CHI_MLU, predict(model_curls, test_data, allow.new.levels = TRUE))
  n = n +1
}

rmse_curls <- data.frame(rmsetest, rmsetrain)
mean(rmse_curls$rmsetest)
#[1] 0.4777694

#Quadratic interaction with gender
n=1
rmsetrain = NULL
rmsetest = NULL

for (f in folds){
  train_data <- subset(data, !(ID %in% f))
  test_data <- subset(data, ID %in% f)
  model_g <- lmer(CHI_MLU ~ visit * diagnosis + I(visit^2) + verbalIQ + (1+visit + I(visit^2) | gender), data = train_data,  REML=FALSE)
  rmsetrain[n] = rmse(train_data$CHI_MLU, predict(model_g))
  rmsetest[n] = rmse(test_data$CHI_MLU, predict(model_g, test_data, allow.new.levels = TRUE))
  n = n +1
}

rmse_g <- data.frame(rmsetest, rmsetrain)
mean(rmse_g$rmsetest)
#[1] 0.5822574

#P-hacked model
n=1
rmsetrain = NULL
rmsetest = NULL

for (f in folds){
  train_data <- subset(data, !(ID %in% f))
  test_data <- subset(data, ID %in% f)
  model_p <- lmer(CHI_MLU ~ visit + I(visit^2) + nonverbalIQ + verbalIQ + types_CHI + MOT_MLU + (1+visit + I(visit^2) | ID), data = train_data, REML=FALSE)
  rmsetrain[n] = rmse(train_data$CHI_MLU, predict(model_p))
  rmsetest[n] = rmse(test_data$CHI_MLU, predict(model_p, test_data, allow.new.levels = TRUE))
  n = n +1
}

rmse_p <- data.frame(rmsetest, rmsetrain)
mean(rmse_p$rmsetest)
#[1] 0.460752
```

#### ANSWER:
Using Cross Validation to calculate accuracy of predictions and using the root mean squared error when applying the actual test data, the best predictive model attempted was our (somewhat p-hacked) model from the previous assignment (pseudocode: Child MLU ~ visit + I(visit^2) + nonverbalIQ + verbalIQ + types_CHI + MOT_MLU + (1+visit + I(visit^2) | ID), producing a RMSE of 0.46. Interestingly, this model does not include ados or diagnosis. It also completely ignores the previously posed hypotheses of the investigation. Should one prefer to stay true to the hypotheses, another model (pseudocode: CHI_MLU ~ visit + I(visit^2) + ados1 + verbalIQ + MOT_MLU + types_CHI + (1+visit + I(visit^2) | ID) performs nearly as well with a RMSE of 0.48. 

### Exercise 3) Assessing the single child

Let's get down to business. This new kiddo - Bernie - has entered your clinic. This child has to be assessed according to his group's average and his expected development.

Bernie is one of the six kids in the test dataset, so make sure to extract that child alone for the following analysis.
```{r, include=FALSE}
#Find Bernie before the data was made anonymous
#filter(test, ID == "Bernie")

#filter out Bernie form the data_test dataframe
sanders <- filter(data_test, ados1 == "7", verbalIQ == "28")

```

You want to evaluate:

- how does the child fare in ChildMLU compared to the average TD child at each visit? Define the distance in terms of absolute difference between this Child and the average TD.
(Tip: recreate the equation of the model: Y=Intercept+BetaX1+BetaX2, etc; input the average of the TD group for each parameter in the model as X1, X2, etc.).

- how does the child fare compared to the model predictions at Visit 6? Is the child below or above expectations? (tip: use the predict() function on Bernie's data only and compare the prediction with the actual performance of the child)

```{r, include=FALSE}
detach(package:plyr)
library(dplyr)

# Filtering out TD participants
TD_full <- data %>% 
  filter(diagnosis == "TD")

# As numeric
TD_full$visit <- as.numeric(TD_full$visit)

# Grouping by visit and summarising to means of required variables
TD_average <- TD_full %>% 
  group_by(visit) %>%
  summarise(mean_TD_mlu = mean(CHI_MLU), verbalIQ = mean(verbalIQ), mot_mlu = mean(MOT_MLU), ados = mean(ados1))

#lmer(CHI_MLU ~ visit + I(visit^2) + verbalIQ + MOT_MLU + ados1 + (1 + visit + I(visit^2) | ID)
#summary(zeus)

#Calculate TD ChildMlu using the model 
intercept <- -1.286345
visit <- 0.452264
visit2 <- -0.036199
ados <- 0.002641  
MOT_MLU <- 0.286893
verbalIQ <- 0.059322

TD1 <- intercept + visit*1 + (visit2*1^2) + ados*TD_average$ados[1] + MOT_MLU*TD_average$mot_mlu[1] + verbalIQ*TD_average$verbalIQ[1]

TD2 <- intercept + visit*2 + (visit2*2^2) + ados*TD_average$ados[2] + MOT_MLU*TD_average$mot_mlu[2] + verbalIQ*TD_average$verbalIQ[2]

TD3 <- intercept + visit*3 + (visit2*3^2) + ados*TD_average$ados[3] + MOT_MLU*TD_average$mot_mlu[3] + verbalIQ*TD_average$verbalIQ[3]

TD4 <- intercept + visit*4 + (visit2*4^2) + ados*TD_average$ados[4] + MOT_MLU*TD_average$mot_mlu[4] + verbalIQ*TD_average$verbalIQ[4]

TD5 <- intercept + visit*5 + (visit2*5^2) + ados*TD_average$ados[5] + MOT_MLU*TD_average$mot_mlu[5] + verbalIQ*TD_average$verbalIQ[5]

TD6 <- intercept + visit*6 + (visit2*6^2) + ados*TD_average$ados[6] + MOT_MLU*TD_average$mot_mlu[6] + verbalIQ*TD_average$verbalIQ[6]

TD_bernie <- select(sanders, visit, CHI_MLU)
TD_bernie$TD_avr <- c(TD1, TD2, TD3, TD4, TD5, TD6)
TD_bernie$abs_difference <- TD_bernie$CHI_MLU-TD_bernie$TD_avr
```

```{r}
TD_bernie
```

```{r, include=FALSE}
data_test$ID[data_test$ID == 1] <- 70
data_test$ID[data_test$ID == 2] <- 71
data_test$ID[data_test$ID == 3] <- 72
data_test$ID[data_test$ID == 4] <- 73
data_test$ID[data_test$ID == 5] <- 74
data_test$ID[data_test$ID == 6] <- 75
data <- data[,-1]
data_test <- data_test[,-1]

data_with_all <- rbind(data, data_test)

data_without_bernie = filter(data_with_all, !(ID == "71" & visit == 6))
data_bernie_6 = filter(data_with_all, (ID == "71" & visit == 6))
bernie_model = lmer(CHI_MLU ~ visit + I(visit^2) + verbalIQ + MOT_MLU + ados1 + ( 1 + visit + I(visit^2) | ID), data = data_without_bernie, REML=FALSE)
predict(bernie_model, newdata = data_bernie_6, allow.new.levels = TRUE)
sanders$CHI_MLU[sanders$visit == 6]
```

If comparing to model calculations of MLU at each visit, Bernie's MLU exceeds the average (model prediction) value for TD children consistently at each visit. As can be read from the table, the absolute difference ranges from roughly 0.58 (at visit 1) to 1.23 (at visit 3) in Bernie's favor. 

Bernie performs above expectation at visit 6 when using all other available data. While the model predicts Bernie's MLU to be 3.13 at visit 6 while his actual MLU was measured to be 3.45. Bernie's general performance and the model predictions of the average TD is illustrated in the plot below (red being Bernie, blue being TD).

```{r}
#Plotting Bernies predictions versus actualt performance 
#theme_set(theme_bw())
bernie_plot <- ggplot(data = TD_bernie, aes(x = visit)) + geom_point(aes(y=CHI_MLU), colour="red") + 
  geom_point(aes(y=TD_avr), colour="blue") +
  geom_line(aes(y=CHI_MLU), colour="red", linetype = 2) + 
  geom_line(aes(y=TD_avr), colour="blue", linetype = 2) +
  xlab('Visit') +
  ylab("Mean Lenght of Utterance")
```

```{r}
bernie_plot
```



CODE DUMP:
```{r}
# Calculating the aboslute difference between Bernie and the average TD child
#TD_average$absolute_difference <- TD_average$bernie-TD_average$mean_TD

# Adding column with model predictions for Bernie
# TD_average$prediction <- predict(zeus, sanders)

# Adding column calculating the absolute accuary of the model's predictions
# TD_average$accuracy <- TD_average$prediction-TD_average$bernie

# model_test <- lmer(CHI_MLU ~ visit + I(visit^2) + ados1 + MOT_MLU + verbalIQ + (1+visit + I(visit^2) | ID), data = data_test, REML=F)

#RMSE for basic model
rmse(data$CHI_MLU, predict(basic))
##[1] 0.3380725

rmse(data_test$CHI_MLU, predict(basic, data_test, allow.new.levels = TRUE))
#[1] 1.009162

rmse(data$CHI_MLU, predict(favmodel))
##[1] 0.2826363

#bestmodel with cross
rmse(data_test$CHI_MLU, predict(favmodel, data_test, allow.new.levels = TRUE))
#[1] 0.7188856

# Examine the split plan
#str(split_plan)

#Some weird means of cross-validation that we can't make sense of (but have run):
folds = createFolds(unique(data$ID), 5)

data$basic_pred <- 0
data$favmodel_pred <- 0
data$model3_pred <- 0
data$model4_pred <- 0
data$p_hacked_pred <- 0

for (f in folds) {
  idx = data$ID %in% f
  train = filter(data, !idx)
  test = filter(data, idx)
  
  basic_train = lmer(CHI_MLU ~ visit + diagnosis + (1 + visit | ID), data = train, REML = FALSE)
  
  favmodel_train = lmer(CHI_MLU ~ visit + I(visit^2) + verbalIQ + MOT_MLU + ados1 + ( 1 + visit + I(visit^2) | ID), data = train, REML=FALSE)
  
  model3 = lmer(CHI_MLU ~ visit*diagnosis + I(visit^2) + (1+visit + I(visit^2) | ID), data = train,  REML=FALSE)
  
  model4 = lmer(CHI_MLU ~ visit * diagnosis + I(visit^2) + verbalIQ + (1+visit + I(visit^2) | gender), data = train,  REML=FALSE)
  
  p_hacked = lmer(CHI_MLU ~ visit + I(visit^2) + nonverbalIQ + verbalIQ + types_CHI + MOT_MLU + (1+visit + I(visit^2) | ID), data = data, REML=FALSE)
  
  data$basic_pred[idx] = predict(basic_train, newdata = test, allow.new.levels = TRUE)
  data$favmodel_pred[idx] = predict(favmodel_train, newdata = test, allow.new.levels = TRUE)
  data$model3_pred[idx] = predict(model3, newdata = test, allow.new.levels = TRUE)
  data$model4_pred[idx] = predict(model4, newdata = test, allow.new.levels = TRUE)
  data$p_hacked_pred[idx] = predict(p_hacked, newdata = test, allow.new.levels = TRUE)
}

rmse(data$CHI_MLU, data$basic_pred)
rmse(data$CHI_MLU, data$favmodel_pred)
rmse(data$CHI_MLU, data$model3_pred)
rmse(data$CHI_MLU, data$model4_pred)
rmse(data$CHI_MLU, data$p_hacked_pred)

#[1] 0.9822976
#[1] 0.8524846
#[1] 0.9457857
#[1] 0.8302081
#[1] 0.6600345

----
rmse <- data.frame(rmsetest, rmsetrain)
mean(rmse$traintest)
mean(rmse$rmsetest)

max(rmse$rmsetrain)-min(rmse$rmsetrain)
max(rmse$rmsetest)-min(rmse$rmsetest)

max(rmse_fav$rmsetrain)-min(rmse_fav$rmsetrain)
max(rmse_fav$rmsetest)-min(rmse_fav$rmsetest)
----
###Cross validation
set.seed(22)

# Implement the 5-fold cross-fold plan
split_plan <- kWayCrossValidation(nrow(data), 5, NULL, NULL)

# Run the 5-fold cross validation plan from split_plan on hercules
k <- 5 # Number of folds

#Setting the variable for the predictive values from the CV
data$pred.cv <- 0

#Creating the loop; uses the basic model and the 5-fold cross-validation plan, gives us predicted values in new column
for(i in 1:k) {
  split <- split_plan[[i]]
  model <- hercules
  data$pred.cv[split$app] <- predict(model, newdata = data[split$app, ])
}

#RMSE for basic model
rmse(data$CHI_MLU, predict(hercules))
##[1] 0.3380725

#Rmse of the cross-validation predictions
rmse(data$CHI_MLU, data$pred.cv)
##[1] 0.5026733

# Run the 5-fold cross validation plan from split_plan on our best model
data$pred.cv.best <- 0

for(i in 1:k) {
  split1 <- split_plan[[i]]
  model2 <- zeus
  data$pred.cv.best[split1$app] <- predict(model2, newdata = data[split1$app, ])
}

# Get the rmse of the full model's predictions
rmse(data$CHI_MLU, predict(zeus))
##[1] 0.2775267

# Get the rmse of the cross-validation predictions
rmse(data$CHI_MLU, data$pred.cv.best)
##[1] 0.4574592

# Loop using Nanna's random model
n=1
rmsetrain = NULL
rmsetest = NULL

for (f in folds){
  train_data <- subset(data, !(ID %in% f))
  test_data <- subset(data, ID %in% f)
  model <- lmer(CHI_MLU ~ visit + I(visit^2) + ados1 + verbalIQ + MOT_MLU + types_CHI + (1+visit + I(visit^2)| ID), data = train_data, REML = FALSE)
  rmsetrain[n] = rmse(train_data$CHI_MLU, predict(model))
  rmsetest[n] = rmse(test_data$CHI_MLU, predict(model, test_data, allow.new.levels = TRUE))
  n = n +1
}

rmse(data$CHI_MLU, data$pred.cv.create1)
## [1] 0.3835551

rmse(predict(model.create1, data_test), data_test$CHI_MLU)
## [1] 0.4976607

# Loop using the p-hacking model from ass. 2
for(i in 1:k) {
  split.create <- split_plan[[i]]
  model.create <- lmer(CHI_MLU ~ visit + I(visit^2) + nonverbalIQ + verbalIQ + types_CHI + MOT_MLU + (1+visit + I(visit^2) | ID), data = data[split.create$train, ], REML=FALSE)
  data$pred.cv.create[split.create$app] <- predict(model.create, newdata = data[split.create$app, ])
}

rmse(data$CHI_MLU, data$pred.cv.create)
## [1] 0.3819715

rmse(predict(model.create, data_test), data_test$CHI_MLU)
## [1] 0.4875191

###We want to have our cake and eat it as well 
#It is a good idea to split in 5 parts (why 5 = )

#We want to make a loop For (F in Folds) select(ID in F)
#Take all the kids that a not in the F fold, and that i our train data set. 
#N=1 in the beginning, Rmse(train[n] =
#N = N+1 

ggplot(data = TD_average, aes(x = visit)) + geom_smooth(aes(y=bernie), colour= "#37a2ae", method = "lm", formula = y ~ poly(x, 3), se = F) + 
         geom_smooth(aes(y=prediction), colour= "#b285dc", method = "lm", formula = y ~ poly(x, 3), se = F) +
  xlab('Number of Visit') +
  ylab("Bernie's Mean Lenght of Utterance")

```

```{r, include=FALSE}
#k <- 10 #the number of folds

#Create the folds
folds <- cvFolds(NROW(data), K=k)

#Set up column in dataframe
data$cv.pred <- rep(0,nrow(data))

#The loop
for(i in 1:k){
  train <- data[folds$subsets[folds$which != i], ] #Set the training set
  validation <- data[folds$subsets[folds$which == i], ] #Set the validation set
  newlm <- lmer(CHI_MLU ~ visit + diagnosis + (1 + visit | ID), data = train, REML = F) #Get your new linear model (just fit on the train data)
  newpred <- predict(newlm,newdata=validation) 
  #Get the predicitons for the validation set (from the model just fit on the train data)

  data[folds$subsets[folds$which == i], ]$cv.pred <- newpred #Put the hold out prediction in the data set for later use
}

#How does the basic model perform
rmse(data$cv.pred, data$CHI_MLU)

```

```{r, include=FALSE}
#preparing a column
data$cv.pred_best <- rep(0,nrow(data))

for(i in 1:k){
  train <- data[folds$subsets[folds$which != i], ] #Set the training set
  validation <- data[folds$subsets[folds$which == i], ] #Set the validation set

  newlmer <- lmer(CHI_MLU ~ visit + I(visit^2) + ados1 + MOT_MLU + (1+visit + I(visit^2) | ID), data = train, REML=FALSE)
  #Get your new model (just fit on the train data)
  
  newpred_best <- predict(newlmer, newdata=validation) 
  #Get the predicitons for the validation set (from the model just fit on the train data)

  data[folds$subsets[folds$which == i], ]$cv.pred_best <- newpred_best 
  #Put the hold out prediction in the data set for later use
}

#How does the best model perform
rmse(data$cv.pred_best, data$CHI_MLU)
```

